# Reproducible Research
## Peer Assessment 1

Let's start with loading the data
```{r,cache=TRUE}
library(knitr)
Data<-read.csv("activity.csv",header=T)
Data$date<-as.Date(Data$date,format="%Y-%m-%d")
cleanData<-na.omit(Data)
```

The first question is: What is mean total number of steps taken per day?
Here is a histogram:
```{r,,cache=TRUE}
total<-by(cleanData$steps,cleanData$date,sum)
plot(total~unique(cleanData$date),type="h", main = "Sum of the steps per day", xlab="Date", ylab="Sum of steps", lwd = 7, col="gray")
mean<-mean(total)
median<-median(total)
```
The mean total number of steps taken per day is `r mean`, and the median is `r median`. They are almost equal, which means that the distribution of steps per day is not skewed.

The second question is: What is the average daily activity pattern?
```{r,,cache=TRUE}
intervals<-unique(cleanData$interval)
a<-split(cleanData$steps, intervals)
avg<-sapply(a, mean)
toPlot<-as.data.frame(cbind(intervals,avg), row.names=NULL)
plot(toPlot$avg~toPlot$intervals,type="l", main = "Average daily activity pattern", xlab="Interval", ylab="Sum of steps")
abline(h=mean(toPlot$avg), col="red")
maxInt<-toPlot[toPlot$avg==max(avg), 1]
```
The plot show that the interval with highest average number of steps taken is `r maxInt`. The red line represents average dayly activity for the investigated period of time.

The next task is to fill the missing values.
```{r}
missing<-is.na(Data)
n.missing<-sum(missing)
```
The total number of row that contain missing values are `r n.missing`.
I chose to substitude the missing values with a random data point from the dataset.
```{r,cache=TRUE}
set.seed(1)
random.imp <- function (a){
      missing <- is.na(a)
      n.missing <- sum(missing)
      a.obs <- a[!missing]
      imputed <- a
      imputed[missing] <- sample (a.obs, n.missing, replace=TRUE)
      return (imputed)
      }
imputedSteps<-random.imp(Data$steps)
impData<-as.data.frame(cbind(imputedSteps,Data$date,Data$interval))
names<-colnames(Data)
colnames(impData)<-names
impData$date<-as.Date(impData$date,format="%Y-%m-%d",origin="1970-01-01")
Total<-by(impData$steps,impData$date,sum)
plot(Total~unique(impData$date),type="h", main = "Total number of steps per day", xlab="Date", ylab="Sum of steps", lwd = 7, col="gray")
Mean<-mean(Total)
Median<-median(Total)
```
The above plot is a copy of the firt one, but here missing values were imputed. As a result the dayly mean is `r Mean` and the dayly median is `r Median`. As a result the mean is almost the same(the diference is in the 5th decimal space), while the median has stayed the same.

The last question is: Are there differences in activity patterns between weekdays and weekends?
First we need to add a weekday/weekend filter. Then create a plot of the average number of sterps per 5-minute interval across weekends and weekdays.
```{r, cache=TRUE}
week<-weekdays(impData$date)
weekFilter<-ifelse(week %in% c("Saurday", "Sunday"), "weekend", "weekday")
LastData<-as.data.frame(cbind(impData,weekFilter))
wd<-subset(LastData, LastData$weekFilter=="weekday")
we<-subset(LastData, LastData$weekFilter=="weekend")
wdS<-split(wd$steps, intervals)
avgWD<-sapply(wdS, mean)
weS<-split(we$steps, intervals)
avgWE<-sapply(weS, mean)
final<-as.data.frame(cbind(intervals,avgWD, avgWE), row.names=NULL)
colnames(final)<-c("interval","weekday","weekend")
par(mfcol=c(2, 1))
plot(final$weekday~final$interval, type="l",main = "Average daily activity pattern: Weekdays", xlab="Interval", ylab="Average number of steps" )
abline(h=mean(final$weekday), col="red")
plot(final$weekend~final$interval, type="l",,main = "Average daily activity pattern: Weekend", xlab="Interval", ylab="Average number of steps" )
abline(h=mean(final$weekend), col="red")
```

The mean for each plot is represented with a red line. Not only that it lies higher for weekends than for weekdays, but other differences could also be observed.
```{r }
knit2html("activity.Rmd")
```